# Reading effluent data from EPA ECHO
wqTools includes two functions to read data from EPA echo. readECHO_fac is used to read facility locations and related metadata. readECHO_ec is used 
to read facility effluent chart data (i.e. water quality sample data). Note that this is just data reported by facilities to ECHO per permit requirements. 
Individual facilities may sample more frequently or for additional parameters.

```{r}
library(wqTools)
library(leaflet)
library(plotly)
```

## Read facility locations
### State-wide facilities
```{r}
ut_fac=readECHO_fac(p_st="ut", p_act="y")
```

### By selected permit IDs
```{r}
gsl_potws=readECHO_fac(p_pid=c("UT0021725","UT0021628","UT0021636","UT0020974","UT0021741","UT0021911","UT0021326","UT0020931","UT0020311","UT0020303","UT0024392","UT0024384"))
buildMap(fac=gsl_potws)
```

## Read facility discharge data
All arguments to readECHO_ec are optional except p_id. At least one p_id must be specified, but readECHO_ec is designed to query data from multiple permit IDs at once. You can also use the [ECHO web services query builder](https://echo.epa.gov/tools/web-services/effluent-charts#/Effluent%20Charts/get_eff_rest_services_download_effluent_chart) 
to read EC data manually, but this is limited to one permit ID at a time.  
Also note that if start_date and end_date are not specified, the service will return the last three years of data. However, older data is available for many facilities.

gsl_ec=readECHO_ec(p_id=c("UT0021725","UT0021628","UT0021636","UT0020974","UT0021741","UT0021911","UT0021326","UT0020931","UT0020311","UT0020303","UT0024392","UT0024384"),
				   start_date="01/01/2000", end_date="09/01/2019", parameter_code=c("50050"))

table(gsl_ec$statistical_base_short_desc)

knitr::kable(with(gsl_ec, {table(npdes_id, statistical_base_short_desc)}))

knitr::kable(with(gsl_ec, {table(npdes_id, standard_unit_desc)}))

gsl_ec$date=as.Date(gsl_ec$monitoring_period_end_date, format='%m/%d/%Y')
gsl_ec$yday=as.POSIXlt(gsl_ec$date)$yday

UT0021725=subset(gsl_ec, npdes_id=='UT0021725')
plot_ly(data=UT0021725, type='scatter') %>%
	add_markers(x=~as.Date(date), mode='markers', y=~dmr_value_standard_units, color=~statistical_base_short_desc)

plot_ly(data=UT0021725, type='scatter') %>%
	add_markers(x=~yday, mode='markers', y=~dmr_value_standard_units, color=~statistical_base_short_desc)








