---
title: "WQP Data Dashboard Report"
output: html_document
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
  spatial_sel_type: NA
  sel_polys: NA
  date_range: NA
  fill_nd: NA
  filters: NA
  multi_site_ts: NA
  multi_site_bp: NA
  conc_map: NA
  report_notes: NA
---


<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r, echo=F}
options(warn=-1)
tags$style(type = "text/css", "html, body {width:100%;height:100%}",
	".leaflet .legend i{
	border-radius: 50%;
	width: 10px;
	height: 10px;
	margin-top: 4px;
	}"
)
```


# Data selections
Data inlcuded in these analyses were selected from:  
`r (params$spatial_sel_type)`(s):  `r params$sel_polys`  
  
These include data from `r params$date_range[1]` to `r params$date_range[2]`.  

<br>

# Data processing
### Non-detect procedure
Non-detect values in these figures are plotted as `r tolower(params$fill_nd)`.

### Filters applied
The following additional filters were applied to the data prior to generating plots:
```{r, echo=F}
params$filters
```

<br>

# Figures
### Multi-site time series
```{r, fig.width=12, echo=F}
params$multi_site_ts
```

### Multi-site boxplot
```{r, fig.width=12, echo=F}
params$multi_site_bp
```

### Concentration map
```{r, fig.width=9, fig.height=5, echo=F}
params$conc_map
```

<br>

# User notes and comments:

`r as.character(params$report_notes)`

<br>
<br>
<br>
<br>
<br>
<br>
<br>

